---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

[![Travis build status](https://travis-ci.org/venelin/BinaryPoissonUsingSPLITT.svg?branch=master)](https://travis-ci.org/venelin/BinaryPoissonUsingSPLITT)
[![Coverage status](https://codecov.io/gh/venelin/BinaryPoissonUsingSPLITT/branch/master/graph/badge.svg)](https://codecov.io/github/venelin/BinaryPoissonUsingSPLITT?branch=master)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/BinaryPoissonUsingSPLITT?color=blue)](http://cran.r-project.org/web/packages/BinaryPoissonUsingSPLITT)
[![Downloads](http://cranlogs.r-pkg.org/badges/BinaryPoissonUsingSPLITT?color=blue)](http://cran.rstudio.com/package=BinaryPoissonUsingSPLITT)


# BinaryPoissonUsingSPLITT

Likelihood calculation of the binary Poisson model using the SPLITT library. This package implementats the example of a discrete model from the article 'Parallel Likelihood Calculation for Phylogenetic Comparative Models: the 'SPLITT' C++ Library'. A preprint of the article is available from https://doi.org/10.1101/235739. The function 'MiniBenchmark' allows to compare the calculation times for different tree sizes.  See [this guide](https://venelin.github.io/SPLITT/articles/SPLITTRcppModules.html) for a tutorial. 

## Installation

You can install the released version of BinaryPoissonUsingSPLITT from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("BinaryPoissonUsingSPLITT")
```

## Examples

* Calculating the likelihood of the PMM for a given tree, data and model parameters.
```{r}
library(ape)
library(BinaryPoissonUsingSPLITT)

N <- 100

x <- 1
q01 <- 0.02
q10 <- 0.8

tree <- rtree(N)
x <- sample(c(0, 1), size = N, replace = TRUE)

cat("logLikelihood using R:", BinaryPoissonModelLogLik(x, tree, 1, 0.02, 0.8), "\n")
cat("logLikelihood using R:", BinaryPoissonModelLogLikCpp(x, tree, 1, 0.02, 0.8), "\n")
```

* Performing a benchmark to measure the likelihood calculation times using different parallelization strategies:

```{r}
# N specifies the size of the phylogenetic tree. 
# Ntests specifies the number of executions in one time measurement 
# (the more the better, but also slower).
MiniBenchmark(N = 100, Ntests = 100)
MiniBenchmark(N = 1000, Ntests = 10)
MiniBenchmark(N = 10000, Ntests = 10)
```
